% Throughout my Ph.D. program and the development of this dissertation, I have received lots of supports and motivation. I especially appreciate valuable motivations from various sources and feel that people value my ability more than what I have.

First and foremost, I would like to thank my academic advisor, Dr. Tommy Dang. Your expertise and expectation are invaluable in guiding me throughout the processes of doing research and publishing work. Your ambition and insightful feedback pushed me out of my own "safety zone" and develop. I would not be able to achieve what I have today without your guidance and supports.

I would particularly like to thank Dr. David C. Weindorf. You are my great source of motivation. You always encourage me to move further from what I have. You value what I offer for our collaboration a lot more than what I have done. Your great collaboration network plays a crucial part in my dissertation development and my achievements during my Ph.D.

I want to thank my collaborators, Dr. Matthew G. Siebecker and Ms. Cynthia M. Jordan. The multivariate proximal sensor data acquired from soil cores and the feedback as the domain experts (soil scientists) provide significant insights for the development of this dissertation. They significantly contribute to the practical uses of the results of this dissertation.

I would also like to thank Dr. Susan A. Mengel particularly. The knowledge that I learned from your Reinforcement Learning class contributed significantly to this dissertation. Thank you for the expertise feedback during this class that helped me experiment with different practical reinforcement learning algorithms and set a foundation for my research in this area.

% I highly appreciate Dr. Rattikorn Hewett as a department chair with excellent supervision regarding research activities and knows well about graduate students in the department. You provided me a perfect opportunity to contribute my service to the department. Your guidance gave me great directions during my job application.

I would like to thank all my dissertation committee members and the graduate dean's representative: Dr. Tommy Dang, Dr. David C. Weindorf, Dr. Matthew G. Siebecker, Dr. Susan A. Mengel, and Dr. Victor Sheng, who all gave valuable comments and suggestions to improve this dissertation. 

I want to acknowledge my colleagues from Interactive Data Visualization Lab. It is fortunate for me to work with you all. I particularly thank my friend, Vinh T. Nguyen. You helped me a lot both in academics and in life during the start of my Ph.D. program.

Last but not least, I greatly appreciate the unconditional support and love that I always have from my parents, parents-in-law, wife, and two little daughters. My parents and parents-in-law always believe in me and put all their hopes for me; my wife gives away her career to support mine; my daughters are always there to give me big hugs every day and show great admiration through every single one of my achievements during my Ph.D. program.